Determining the alternatives for scalar implicature

### 2016-01-24 Paper/Paper.Rmd
minor spelling tweaks, commented out library(langcog) - not compatibile with R version 3.2.3 - only impacts solorazied ggplot coloring

```{r, libraries}
library(ggplot2)
library(knitr)
library(rjson)
library(dplyr)
library(tidyr)
library(gridExtra)
library(jsonlite)
library(png)
library(grid)
library(langcog)
```

@@

Successful communication requires listeners to make pragmatic inferences that go beyond the literal semantic content of speakers' utterances. For example, listeners  commonly enrich the meaning of the scalar item "some" to "some but not all" in sentences like "Bob ate some of the cookies" [@grice1975;@horn1984;@levinson2000]. These inferences, called _scalar implicatures_, have been an important test case for understnading pragmatic inferences more generally. A Gricean account of this phenomena assumes listeners reason about the meaning the speaker intended by incorporating knowledge about a) alternative scalar items a speaker could have used (such as "all") and b) the relative informativity of using such alternatives [@grice1975]. According to this account, a listener will infer that the speaker must have intended that Bob did not eat "all" the cookies because it would have been *underinformative* for the speaker to use "some" when "all" could have been used.

@@

@franke2014 used an RSA-style model to assess what alternatives a speaker would need to consider in order to produce the typicality/felicity ratings reported by @degen2015 and @vantiel2014. In order to do this, @franke2014's model assigned weights to a set of alternative numerical expressions. Surprisingly, along with weighting "one" highly (a conclusion that was suppored by the empirical work),the best-fitting model assigned substantial weight to "none" as an alternative. This finding was especially surprising considering the emphasis of standard theories on scalar items that stand in entailment relationships with one another (e.g. "one" entails "some" even if it is not classically considered to be part of the scale scale).

@@

This model of the pragmatic speaker is consistent with a speaker who choses words to maximize the utility of an utterance in context [@frank2012], where utility is operationalized as the informativity of a particular utterance (surprisal) minus a cost:

@@

The two-alternatives condition included only the scalar pairs from @vantiel2014. The four-alternatives condition included the two scalar items plus the top two alternatives generated for each scalar family by participants in Experiment 2. The five-alternatives condition included the four previous items plus one more "neutral" item chosen from those alternatives generated in Experiment 2.

@@

ggplot(speakers, 
       aes(x = stars, y = speaker.p, col = degree, lty = expt)) + 
  geom_line() + 
  facet_grid(.~ scale) + 
  guides(colour = guide_legend(direction = "horizontal", nrow = 1, 
                               title = "Scalar Item"), 
         linetype = guide_legend(direction = "horizontal", nrow = 1, 
                                 title = "Experiment")) + 
  xlab("Star Rating") +
  ylab("Proportion Judgments") + 
  scale_colour_solarized() + 
  theme_bw() + 
  theme(legend.position = "bottom",
        legend.margin = unit(.02,"in"),
        legend.key = element_blank(), 
        legend.key.size = unit(.1, "in"))
```

@@

To elicit empirical alternatives for the scales we used in Experiment 1, we adopted a modified cloze task inspired by Experiment 2 of @vantiel2014.

@@

Participants were presented a target scalar term from our original entailment set embedded in a sentence such as, "In a recent restaurant review someone said they thought they the food was \____," with a target scalar presented in the "\____." Participants were then asked to generate plausible alternatives by responding to the question, "If they'd felt differently about the food, what other words could they have used instead of \_____?" They were prompted to generate three unique alternatives.

>>>>

```{r, libraries}
library(ggplot2)
library(knitr)
library(rjson)
library(dplyr)
library(tidyr)
library(gridExtra)
library(jsonlite)
library(png)
library(grid)
# library(langcog)
```

@@

Successful communication requires listeners to make pragmatic inferences that go beyond the literal semantic content of speakers' utterances. For example, listeners  commonly enrich the meaning of the scalar item "some" to "some but not all" in sentences like "Bob ate some of the cookies" [@grice1975;@horn1984;@levinson2000]. These inferences, called _scalar implicatures_, have been an important test case for understanding pragmatic inferences more generally. A Gricean account of this phenomena assumes listeners reason about the meaning the speaker intended by incorporating knowledge about a) alternative scalar items a speaker could have used (such as "all") and b) the relative informativity of using such alternatives [@grice1975]. According to this account, a listener will infer that the speaker must have intended that Bob did not eat "all" the cookies because it would have been *underinformative* for the speaker to use "some" when "all" could have been used.

@@

@franke2014 used an RSA-style model to assess what alternatives a speaker would need to consider in order to produce the typicality/felicity ratings reported by @degen2015 and @vantiel2014. In order to do this, @franke2014's model assigned weights to a set of alternative numerical expressions. Surprisingly, along with weighting "one" highly (a conclusion that was suppored by the empirical work), the best-fitting model assigned substantial weight to "none" as an alternative. This finding was especially surprising considering the emphasis of standard theories on scalar items that stand in entailment relationships with one another (e.g. "one" entails "some" even if it is not classically considered to be part of the scale).

@@

This model of the pragmatic speaker is consistent with a speaker who chooses words to maximize the utility of an utterance in context [@frank2012], where utility is operationalized as the informativity of a particular utterance (surprisal) minus a cost:

@@

The two-alternatives condition included only the scalar pairs from @vantiel2014. The four-alternatives condition included the two scalar items plus the top two alternatives generated for each scalar family by participants in Experiment 2. The five-alternatives condition included the four previous items plus one more "neutral" item chosen from those alternatives generated in Experiment 2. See Figure 1 for the complete list of alternatives used in each condition.

@@

ggplot(speakers, 
       aes(x = stars, y = speaker.p, col = degree, lty = expt)) + 
  geom_line() + 
  facet_grid(.~ scale) + 
  guides(colour = guide_legend(direction = "horizontal", nrow = 1, 
                               title = "Scalar Item"), 
         linetype = guide_legend(direction = "horizontal", nrow = 1, 
                                 title = "Experiment")) + 
  xlab("Star Rating") +
  ylab("Proportion Judgments") + 
  # scale_colour_solarized() + 
  theme_bw() + 
  theme(legend.position = "bottom",
        legend.margin = unit(.02,"in"),
        legend.key = element_blank(), 
        legend.key.size = unit(.1, "in"))
```

@@

To elicit empirical alternatives for the scales we used in Experiment 1, we adopted a modified cloze task inspired by Experiment 3 of @vantiel2014.

@@

Participants were presented a target scalar term from our original entailment set embedded in a sentence such as, "In a recent restaurant review someone said they thought they the food was \____," with a target scalar presented in the blank. Participants were then asked to generate plausible alternatives by responding to the question, "If they'd felt differently about the food, what other words could they have used instead of \_____?" They were prompted to generate three unique alternatives.

### 2016-01-24 Paper/Paper.tex
minor spelling tweaks, commented out library(langcog) - not compatibile with R version 3.2.3 - only impacts solorazied ggplot coloring

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

@@

Successful communication requires listeners to make pragmatic inferences
that go beyond the literal semantic content of speakers' utterances. For
example, listeners commonly enrich the meaning of the scalar item
``some'' to ``some but not all'' in sentences like ``Bob ate some of the
cookies'' (Grice, 1975; Horn, 1984; Levinson, 2000). These inferences,
called \emph{scalar implicatures}, have been an important test case for
understnading pragmatic inferences more generally. A Gricean account of
this phenomena assumes listeners reason about the meaning the speaker
intended by incorporating knowledge about a) alternative scalar items a
speaker could have used (such as ``all'') and b) the relative
informativity of using such alternatives (Grice, 1975). According to
this account, a listener will infer that the speaker must have intended
that Bob did not eat ``all'' the cookies because it would have been
\emph{underinformative} for the speaker to use ``some'' when ``all''
could have been used.

@@

Franke (2014) used an RSA-style model to assess what alternatives a
speaker would need to consider in order to produce the
typicality/felicity ratings reported by Degen \& Tanenhaus (2015) and
Tiel (2014). In order to do this, Franke (2014)'s model assigned weights
to a set of alternative numerical expressions. Surprisingly, along with
weighting ``one'' highly (a conclusion that was suppored by the
empirical work),the best-fitting model assigned substantial weight to
``none'' as an alternative. This finding was especially surprising
considering the emphasis of standard theories on scalar items that stand
in entailment relationships with one another (e.g. ``one'' entails
``some'' even if it is not classically considered to be part of the
scale scale).

@@

This model of the pragmatic speaker is consistent with a speaker who
choses words to maximize the utility of an utterance in context (Frank
\& Goodman, 2012), where utility is operationalized as the informativity
of a particular utterance (surprisal) minus a cost:

@@

\begin{CodeChunk}
\captionsetup{width=0.8\textwidth}\begin{figure*}[t]

@@

The two-alternatives condition included only the scalar pairs from Tiel
(2014). The four-alternatives condition included the two scalar items
plus the top two alternatives generated for each scalar family by
participants in Experiment 2. The five-alternatives condition included
the four previous items plus one more ``neutral'' item chosen from those
alternatives generated in Experiment 2.

@@

To elicit empirical alternatives for the scales we used in Experiment 1,
we adopted a modified cloze task inspired by Experiment 2 of Tiel
(2014).

@@

Participants were presented a target scalar term from our original
entailment set embedded in a sentence such as, ``In a recent restaurant
review someone said they thought they the food was \_\_\_\_," with a
target scalar presented in the ``\_\_\_\_." Participants were then asked
to generate plausible alternatives by responding to the question, ``If
they'd felt differently about the food, what other words could they have
used instead of \_\_\_\_\_?'' They were prompted to generate three
unique alternatives.

@@

\begin{CodeChunk}
\captionsetup{width=0.8\textwidth}\begin{figure}[t]

>>>>

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}

@@

Successful communication requires listeners to make pragmatic inferences
that go beyond the literal semantic content of speakers' utterances. For
example, listeners commonly enrich the meaning of the scalar item
``some'' to ``some but not all'' in sentences like ``Bob ate some of the
cookies'' (Grice, 1975; Horn, 1984; Levinson, 2000). These inferences,
called \emph{scalar implicatures}, have been an important test case for
understanding pragmatic inferences more generally. A Gricean account of
this phenomena assumes listeners reason about the meaning the speaker
intended by incorporating knowledge about a) alternative scalar items a
speaker could have used (such as ``all'') and b) the relative
informativity of using such alternatives (Grice, 1975). According to
this account, a listener will infer that the speaker must have intended
that Bob did not eat ``all'' the cookies because it would have been
\emph{underinformative} for the speaker to use ``some'' when ``all''
could have been used.

@@

Franke (2014) used an RSA-style model to assess what alternatives a
speaker would need to consider in order to produce the
typicality/felicity ratings reported by Degen \& Tanenhaus (2015) and
Tiel (2014). In order to do this, Franke (2014)'s model assigned weights
to a set of alternative numerical expressions. Surprisingly, along with
weighting ``one'' highly (a conclusion that was suppored by the
empirical work), the best-fitting model assigned substantial weight to
``none'' as an alternative. This finding was especially surprising
considering the emphasis of standard theories on scalar items that stand
in entailment relationships with one another (e.g. ``one'' entails
``some'' even if it is not classically considered to be part of the
scale).

@@

This model of the pragmatic speaker is consistent with a speaker who
chooses words to maximize the utility of an utterance in context (Frank
\& Goodman, 2012), where utility is operationalized as the informativity
of a particular utterance (surprisal) minus a cost:

@@

\begin{CodeChunk}
\begin{figure*}[t]

@@

The two-alternatives condition included only the scalar pairs from Tiel
(2014). The four-alternatives condition included the two scalar items
plus the top two alternatives generated for each scalar family by
participants in Experiment 2. The five-alternatives condition included
the four previous items plus one more ``neutral'' item chosen from those
alternatives generated in Experiment 2. See Figure 1 for the complete
list of alternatives used in each condition.

@@

To elicit empirical alternatives for the scales we used in Experiment 1,
we adopted a modified cloze task inspired by Experiment 3 of Tiel
(2014).

@@

Participants were presented a target scalar term from our original
entailment set embedded in a sentence such as, ``In a recent restaurant
review someone said they thought they the food was \_\_\_\_," with a
target scalar presented in the blank. Participants were then asked to
generate plausible alternatives by responding to the question, ``If
they'd felt differently about the food, what other words could they have
used instead of \_\_\_\_\_?'' They were prompted to generate three
unique alternatives.

@@

\begin{CodeChunk}
\begin{figure}[t]
